\chapter{Nice Things from C++11}

Ah, \cpp11, the long awaited update of the language. There was so much
hope in it! Just imagine, back in the day the language had no
automatic dynamic memory management, and not even a hash table in the
standard library… in the year 2000!

At that time we were trying to compensate for the lack of features
with Boost, eagerly waiting for the ever postponed upcoming \cpp 0x
that would change everything. We were going to have modules! And
concepts too!

Looking back into these years, I am quite happy that the committee
dropped the most complex features to finally publish a new version,
because most of the parts that were ready are quite awesome.

\section{At the Language Level}
\input{parts/11/range-based-for-loops}
\input{parts/11/nullptr}
\input{parts/11/scoped-enumerations}
\input{parts/11/constexpr}
\input{parts/11/uniform-initialization}
\input{parts/11/static-assert}

\input{parts/11/lambdas}
\input{parts/11/decltype}
\input{parts/11/auto}
\input{parts/11/rvalue-reference}

\section{Move Semantics}
\input{parts/11/move}

\section{Template-Related Features}
\input{parts/11/template-aliases}
\input{parts/11/extern-template}
\input{parts/11/variadic-template}

\section{Objects and Classes}
\input{parts/11/delegated-constructors}
\input{parts/11/deleted-constructors}
\input{parts/11/defaulted-constructors}
\input{parts/11/override-and-final}
\input{parts/11/using-parent-function}

\section{In the Standard Library}
\input{parts/11/free-begin}
\input{parts/11/prev-next}
\input{parts/11/array}
\input{parts/11/random}
\input{parts/11/smart-pointers}
\input{parts/11/initializer-list}
\input{parts/11/thread}
\input{parts/11/tuple}
\input{parts/11/bind}
\input{parts/11/reference-wrapper}
\input{parts/11/function}
\input{parts/11/hash-table}
\input{parts/11/type-traits}

% chrono is missing

\section{Miscellaneous}
\input{parts/11/miscellaneous}

% member initialization
% inline namespaces
