%-------------------------------------------------------------------------------
\chapter{Preface}
\renewcommand*\thesection{\arabic{section}}

%-------------------------------------------------------------------------------
\section{About this Book}

Once upon a time, as I was working on a large project with others
\cpp{} programmers, I was asked to set up a series of talks about the
language and especially about what has changed since the arrival of
\cpp11. It was in 2020.

So I started to write some slides with what seemed to be the key
features from \cpp11, and quite soon I had to face the truth: it is a
lot of content, and there was three additional major updates in the
language that should be covered too.

In the end I did not do the talks. However, I kept working on the
slides, until eventually I decided to switch the format. It is
probably too much material for a talk, but what about a small book?

Hence this document.

\bigskip

The goal is to list many, if not all, essential features introduced in
the \cpp{} language since its first well-known deep update, known as
\cpp11, up to the most recent version of the standard, which is \cpp20
by the time I am writing this.

These features are for the most part presented following a format
where the pre-\cpp11 way is reminded to the reader, with a short
explanation of why it may have been problematic or inefficient, then
the new way of doing things is presented.

Some parts of the language are silenced, mostly the ones for which I
don't know much, other parts are more thoroughly presented. In any
case, this book won't go into the details and subtleties of any
feature, nor into compiler-specific stuff. The reason being mostly
time (as far as I can tell I have a limited amount of that in my life)
and space (the book is already large enough). The reader is invited to
satisfy his curiosity and complete is knowledge by reading other
material. For example, the website \url{https://en.cppreference.com}
has everything we need to know about any feature of the language.

Be advised that some critics may suddenly appear in these pages, about
the language or the programmers. Keep in mind that those are personal
opinions and may change suddenly!

Finally, a basic knowledge of the language is preferable for the
reader to enjoy this book, as some notions will be used without being
explained.

\bigskip

While this book is about \cpp, one should remember that \cpp{} itself
is just a tool in the programmer's toolbox. If you are focusing on
learning \cpp{} to become a programmer, a good programmer, I would
suggest to rethink your plan and learn programming on a larger scale:
computer architecture, algorithms, data structures, project
management, packaging, dependency management, coding style, reviews,
testing…  There are many aspects to be familiar with in the daily life
of a programmer, keep some place for them.

As a good starting point, every programmer should read Code Complete
\cite{code-complete}. This book goes in detail in all aspects of
software development, backed up by data coming from over decades of
real-life projects, so if you read it you will also gain part of the
knowledge from these people who tried, failed, and succeeded before
you.

Clean Code \cite{clean-code} is a good second book to read, even
though I would not approve all suggestions. For example, it pushes for
intensive factorization and the use of object-oriented programming
everywhere, which are rather things I have painfully learnt use
parsimoniously. Still, the book is a reference in software
development, so you should read it at least to make yourself an
opinion and to know what is going on in the business.

\bigskip

Finally, remember that if reading is acquiring the experience of
others, practicing is building our own experience. So I strongly
recommend to find or start a side project, maybe even a rewrite of
existing tools, just to try and get a grasp of the potential underlying
complexity.

%-------------------------------------------------------------------------------
\section{About the Author}

Should one take this book's content at face? Is the author legit? It
is normal to question the legitimacy of who pretend to give advice.
So in order to help you gauging the credibility of this book, I think
it is important to tell a bit about me.

\bigskip

First of all, I read a lot of code. I read code almost everyday, on
GitHub, on blogs, on StackOverflow, Reddit, and on other forums. I
read code written by me or others, from my personal projects, from my
employer, or from random projects I find on the Internet. All this
code is displayed either on my laptop, or in a terminal connected to a
remote server, or on my phone. Reading is undoubtedly the main part of
my programming activities.

On the productive side, I write code as a hobby since 1994, and
professionally since 2005. I have at least half a million of \cpp{}
behind me, just counting the lines of code that survived in past
projects I could find.

I also have coded a lot of Bash, a good share of Java, a bit of HTML
and a bit of JavaScript. Additionally, in a more anecdotal way, I
coded some C\# programs, some ActionScript, some Pascal and Delphi
ones, a small compiler in Eiffel, some pet projects in Visual Basic
too, BASIC a long time ago, on a Commodore 128 and later under DOS. I
also did a bit of Objective-C.

Let's face it though, a good share of this code was crap.

\bigskip

Some code did end up well nonetheless. One project I am proud of is a
mobile game written in \cpp, which was played by more than 500'000
people every day during more than three years. Aside from that, I also
took part in projects that were struggling to start and brought them
into a viable product. So I guess I made stuff that does not suck.

When I code I tend to think about long term and architecture. I try
not to take any shortcut and to answer the problem without attempting
to solve the future. I code in small boxes, many, with the intent that
they can be broken, removed, replaced, without changing everything. It
wasn't always like that but that's how I work today.

Finally, I am certainly not the type to rush for the new thing. I like
tools and practices that have been well tested, so you probably won't
hear me telling you to use this new thing from \cpp{42} because it's
new and it will show that you are modern and blah blah blah.

\bigskip

Convinced? Anyway, I hope you will find something useful in this book :)

%-------------------------------------------------------------------------------
\section{License}
This work of art is licensed under a Creative Commons
Attribution-ShareAlike 4.0 International License. See
Appendix~\ref{license} for the full license text.

\marginheader{This one.}%
%
The header file icon displayed in the margins, like the one on the
side of this paragraph, is based on the New Document icon from the
Tango Icon Theme \cite{tango-icon-theme}. The original icon is in the
public domain and in order to respect the intent of the source the
variation made for this book, as in the SVG file available in the
repository containing the sources of this book, is also released in
the public domain.

\renewcommand*\thesection{\arabic{chapter}.\arabic{section}}
