%-------------------------------------------------------------------------------
\subsection{Deleted Constructors}

\problemtitle

Delegating constructors is a nice feature, but wat about totally
removing a constructor?

Consider the class below:

\begin{lstlisting}
struct scoped_listener
{
  scoped_listener(dispatcher& d, callback c)
    : m_dispatcher(&d)
  {
    m_id = m_dispatcher->connect(c);
  }

  ~scoped_listener()
  {
    m_dispatcher->disconnect(m_id)
  }

private:
  dispatcher* m_dispatcher;
  int m_id;
};
\end{lstlisting}

Creating copies of \code{scoped\_listener} does not make any sense, as
all copies would share the same \code{m\_id} and will thus trigger the
same call to \code{disconnect(int)} when destructed. See for example
its usage below:

\begin{lstlisting}
void foo()
{
  /* ... */
  scoped_listener listener(d, c1);
  scoped_listener copy(listener);
  scoped_listener other(d, c2);

  {
    // This assignment does not call disconnect(c2).
    other = listener;
    // disconnect(c1) is called here,
    // when other goes out of scope.
  }
  // disconnect(c1) is called twice here: in the
  // destruction of listener and copy.
}
\end{lstlisting}

One would typically want to forbid copies of \code{scoped\_listener} by
disabling its copy constructor.

Before \cpp11, one solution we would find here and there was to
declare the copy constructor and assignment operator as private. The
problem was that it was still available for the class and its
friends. So the programmer would then either implement an always
failing body for this constructor, emitting an error at run time, or
would just not implement the constructor, thus triggering an error at
link time.

These solutions were kind of weak, in the sense that the error, if
any, was presented quite late for the programmer, and with a not
obvious explanation.

\solutiontitle

Now, starting with \cpp11, the constructor and operators can be
explicitly deleted:

\begin{lstlisting}
struct scoped_listener
{
  scoped_listener(const scoped_listener&)@\emcode{ = delete}@;
  scoped_listener& operator=(const scoped_listener&)@\emcode{ = delete}@;

  scoped_listener(dispatcher& d, callback c)
    : m_dispatcher(&d)
  {
    m_id = m_dispatcher->connect(c);
  }

  ~scoped_listener()
  {
    m_dispatcher->disconnect(m_id)
  }

private:
  dispatcher* m_dispatcher;
  int m_id;
};
\end{lstlisting}

Using a deleted function will trigger a clear error from the compiler
when the call is encountered.
